CUDA-accelerated genetic feedforward-ANN training for data mining

### 2010-07-10 doc/HPCS-proceedings-paper.tex
Add clarification to computation of ANN values
git-svn-id: http://svn.scesoc.ca/evolution/trunk@127 fee98b4c-d1c3-4d34-9525-a468159e4a48

% ----------------------------------------------------------------------------
\subsection{Computation of ANN Output Values} \label{ANNcomp}
% ----------------------------------------------------------------------------
ANN output values are computed by reading the training instances as real-valued feature vectors and applying the appropriate operations represented by the network topology and parameters. Feature data are organized to efficiently use memory bandwidth. ANN output calculation is particularly well-suited to the CUDA architecture because it has a very high ratio of mathematical operations to memory accesses (160 multiplications, 4 exponentiations and 80 bytes of memory accesses for our ANN topology). In addition, each output value depends only on the current ANN parameters and one training instance's feature values. Therefore, parallelization of this step scales particularly well with hardware capabilities.

>>>>

% ----------------------------------------------------------------------------
\subsection{Computation of ANN Output Values} \label{ANNcomp}
% ----------------------------------------------------------------------------
ANN output values are computed by reading the training instances as real-valued feature vectors, applying them to the input nodes of the network (see Figure~\ref{fig:ann-chromosome}), and storing the output of the network. Specifically, each feature value is first scaled by the corresponding input-hidden edges. These values are then propagated to all hidden nodes, which apply a radial basis function. Finally, the hidden node outputs are scaled by the hidden-output edges and summed at the output node. A single real value is output from the network and stored in GPU memory. Parameters for edge weights and radial basis functions are given in the chromosome of each candidate ANN.

Feature data are organized to efficiently use memory bandwidth. The requirement for optimal use of CUDA memory bandwidth is that blocks of threads simultaneously access sequential addresses in memory, resulting in a coalesced single-clock cycle access. We store feature data in feature-major order (values for all instances of feature 1, all instances of feature 2, etc.). Therefore, when the thread block reads data for the first feature across different instances, the memory accesses can be coalesced. Similarly, accesses are coalesced for all subsequent features.

ANN output calculation is particularly well-suited to the CUDA architecture because it has a very high ratio of mathematical operations to memory accesses (160 multiplications, 4 exponentiations and 80 bytes of memory accesses for our ANN topology). In addition, each output value depends only on the current ANN parameters and one training instance's feature values. Therefore, parallelization of this step scales particularly well with hardware capabilities.
